---
title: "ST558 Project 2" 
author: "Noel Hilliard"
date: "July 3, 2020"
output:
  rmarkdown::github_document:
    toc: TRUE
params: 
    day: "monday"
---

Project Objective: The goal is to create models for predicting the `shares` variable from the dataset. You will create two models: a linear regression model and a non-linear model (each of your choice). You will use the parameter functionality of markdown to automatically generate an analysis report for each `weekday_is_*` variable (so you’ll end up with seven total outputted documents).


```{r libraries, echo = TRUE, message = FALSE}
library(ggplot2)
library(Hmisc)
library(rmarkdown)
library(tidyverse)
library(corrplot)
library(knitr)
library(caret)

```




```{r rmarkdown params, eval = FALSE, echo = FALSE}

#days of the week
days <- c("monday", "tuesday")

#days <- c("monday", "tuesday", "wednesday",
           #"thursday", "friday", "aturday",
           #"sunday")

#create file names
output_file <- paste0(days, "Analysis.md")

#create a list for each day of the week with the day as parameter
params <- lapply(days, FUN = function(x){list(day = x)})

#put into data frame
#reports <- tibble(days, filename = output_file, params = purrr::map(days, ~list(day = .)))
reports <- tibble(filename = output_file, params)
```

```{r rmarkdown automation, eval = FALSE, echo = FALSE}
apply(reports, MARGIN = 1, FUN = function(x){
  render("README.Rmd", 
         output_file = x[[1]], params = x[[2]])
})
```


# Introduction Online News Popularity Data

# `r capitalize(params$day)` Data

```{r read/filter data, echo = TRUE, message = FALSE}

#read in data
newsData <-read_csv("S:/ST558/Homeworks/Project 2/ST558 Project 2/OnlineNewsPopularity.csv")
newsData$shares_group <- NA

#filter for day of week
weekdayData <- filter(newsData, newsData[[paste0("weekday_is_",params$day)]] == "1")


#split shares into 2 groups
for (i in 1:length(weekdayData$shares)){
  
  if(weekdayData$shares[i] < 1400){
    weekdayData$shares_group[i] <- "below 1400"
  }
  
  if(weekdayData$shares[i] >= 1400){
    weekdayData$shares_group[i] <- "above 1400"
  }
}

weekdayData$shares_group <- as.factor(weekdayData$shares_group)

```

```{r train/test data, echo = TRUE}

#Create test/train data sets from filtered data set
set.seed(1)
train <- sample(1:nrow(weekdayData), size = nrow(weekdayData)*0.7)
test <- dplyr::setdiff(1:nrow(weekdayData), train)
weekdayDataTrain <- weekdayData[train, ]
weekdayDataTest <- weekdayData[test, ]
```

# Summary Statistics

```{r histogram, echo = TRUE}
g <- ggplot(weekdayDataTrain, aes(x = shares))
g + geom_histogram(bins = 100)
```

```{r bar plot, echo = TRUE}

g <- ggplot(data = weekdayDataTrain, aes(x = shares_group))
g + geom_bar() + labs(x = "Shares Group", title = (paste0(capitalize(params$day)," Shares Groups")))
```

```{r shares_group count, echo = TRUE}
table(weekdayDataTrain$shares_group)
```

```{r TrainData stat summary, echo = TRUE}

TrainStatSum <- function(group){
  data <- weekdayDataTrain %>% filter(shares_group == group) %>% 
    select(num_keywords, avg_positive_polarity, num_videos, num_imgs, 
           max_positive_polarity, title_subjectivity, rate_negative_words,
           n_unique_tokens, average_token_length, global_rate_positive_words)
           kable(apply(data, 2, summary, col.names = c("num_keywords", 
           "avg_positive_polarity", "num_videos", "num_imgs", "max_positive_polarity", 
           "title_subjectivity", "rate_negative_words", "n_unique_tokens", "average_token_length",
           "global_rate_positive_words")))
}

```

```{r above 1400, echo = TRUE}
TrainStatSum("above 1400")
```

```{r below 1400, echo = TRUE}
TrainStatSum("below 1400")
```

```{r correlation, echo = TRUE}
correlation <- cor(weekdayDataTrain %>% select(num_keywords, avg_positive_polarity, num_videos, num_imgs, 
           max_positive_polarity, title_subjectivity, rate_negative_words,
           n_unique_tokens, average_token_length, global_rate_positive_words), method = "spearman")

corrplot(correlation, type = "upper", tl.pos = "lt")
```


# Modeling

## Ensemble Model: Bagged Tree

```{r bagged tree , echo = TRUE, message = FALSE}

set.seed(1)
trCtrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

bag_fit <- train(shares_group ~ num_keywords + avg_positive_polarity + num_videos + num_imgs, 
                 data = weekdayDataTrain, method = "treebag", trControl = trCtrl, 
                 preProcess = c("center", "scale"), na.action = na.pass)

bag_fit
```

```{r predict bag tree , echo = TRUE}
bag_pred <- predict(bag_fit, newdata = weekdayDataTest)

head(bag_pred)
```

```{r bag fit info , echo = TRUE}
bag_fitInfo <- tbl_df(data.frame(bag_pred, weekdayDataTest$shares_group, weekdayDataTest$num_keywords, 
                                weekdayDataTest$avg_positive_polarity,
                                 weekdayDataTest$num_videos, weekdayDataTest$num_imgs))

bag_fitInfo
```

```{r bag table , echo = TRUE}
bag_tbl <- table(bag_fitInfo$bag_pred, bag_fitInfo$weekdayDataTest.shares_group)

bag_tbl
```

```{r bag misclass, echo = TRUE}
bag_misClass <- 1 - sum(diag(bag_tbl))/sum(bag_tbl)

bag_misClass
```

## Select Best Multiple Linear Regression Model

```{r mlr, echo = TRUE}
mlrFit1 <- lm(shares ~ num_keywords + avg_positive_polarity + 
                num_videos + num_imgs, data = weekdayDataTrain)

mlrFit2 <- lm(shares ~ max_positive_polarity + title_subjectivity, 
              data = weekdayDataTrain)

mlrFit3 <- lm(shares ~ num_keywords + rate_negative_words,
              data = weekdayDataTrain)

mlrFit4 <- lm(shares ~ n_unique_tokens + average_token_length + 
                 global_rate_positive_words, data = weekdayDataTrain)

```

```{r compare fit, echo = TRUE}

library(MuMIn)

compareFitStats <- function(mlrFit1, mlrFit2, mlrFit3, mlrFit4){
  fitStats <- data.frame(fitStat = c("Adj R Square", "AIC", "BIC"),
  col1 = c(round(summary(mlrFit1)$adj.r.squared, 5), AIC(mlrFit1),
                 BIC(mlrFit1)),
  col2 = c(round(summary(mlrFit2)$adj.r.squared, 5), AIC(mlrFit2),
                 BIC(mlrFit2)),
  col3 = c(round(summary(mlrFit3)$adj.r.squared, 5), AIC(mlrFit3),
                BIC(mlrFit3)),
  col4 = c(round(summary(mlrFit4)$adj.r.squared, 5), AIC(mlrFit4),
                BIC(mlrFit4)))
  
  #put names in data frame
  calls <- as.list(match.call())
  calls[[1]] <- NULL
  names(fitStats[2:5]) <- unlist(calls)
  fitStats
}
```

```{r fit stats, echo = TRUE}
fitStats <- compareFitStats(mlrFit1, mlrFit2, mlrFit3, mlrFit4)
```

```{r compare rmse, echo = TRUE}
mlrFit1_pred <- predict(mlrFit1, newdata = weekdayDataTest)
RMSE(weekdayDataTest$shares, mlrFit1_pred)

mlrFit4_pred <- predict(mlrFit4, newdata = weekdayDataTest)
RMSE(weekdayDataTest$shares, mlrFit4_pred)
```

