---
title: "ST558 Project 2" 
author: "Noel Hilliard"
date: "July 3, 2020"
output:
  rmarkdown::github_document:
    toc: TRUE
---

Project Objective: The goal is to create models for predicting the `shares` variable from the dataset. You will create two models: a linear regression model and a non-linear model (each of your choice). You will use the parameter functionality of markdown to automatically generate an analysis report for each `weekday_is_*` variable (so you’ll end up with seven total outputted documents).

The code below was used to create .md files for each day of the week by calling the READMEtemplate.Rmd file in the `rmarkdown::render` function. The `params` variable created a parameter for each day of the week and the `output_file` variable created a file name for each day of the week.

```{r , echo = TRUE, eval = TRUE, message = FALSE}
library(rmarkdown)
library(tidyverse)
library(Hmisc)

#days of the week
days <- c("monday", "tuesday", "wednesday",
          "thursday", "friday", "saturday",
           "sunday")

#create file names
output_file <- paste0(capitalize(days), "Analysis.md")

#create a list for each day of the week with the day as parameter
params <- lapply(days, FUN = function(x){list(day = x)})

#put into data frame
#reports <- tibble(days, filename = output_file, params = purrr::map(days, ~list(day = .)))
reports <- tibble(filename = output_file, params)
```

```{r , echo = TRUE, message = FALSE, results = "hide"}
apply(reports, MARGIN = 1, FUN = function(x){
  render("READMEtemplate.Rmd", 
         output_file = x[[1]], params = x[[2]])
})
```

# Introduction Online News Popularity Data

The data used for this analysis was obtained from the [UCI Machine Learning Reposititory](https://archive.ics.uci.edu/ml/datasets/Online+News+Popularity). The data set summarizes a set of features about articles published by Mashable over a two year period. The purpose of this analysis is to predict the number of shares in social networks using select features from the data set. Two types of methods will be used to predict the `shares`. The first model discussed is a bagged tree model. For this analysis, the `shares` were divided into two groups (< 1400 and >= 1400) to create a binary classification problem. The second model discussed is linear regression. Both types of models were trained/tuned using the training data set and then then predictions were made using the test data set.

# Analysis

The full data set contained data for all days of the week. 

* The analysis for [Monday is available here.](MondayAnalysis.md)

* The analysis for [Tuesday is available here.](TuesdayAnalysis.md)

* The analysis for [Wednesday is available here.](WednesdayAnalysis.md)

* The analysis for [Thursday is available here.](ThursdayAnalysis.md)

* The analysis for [Friday is available here.](FridayAnalysis.md)

* The analysis for [Saturday is available here.](SaturdayAnalysis.md)

* The analysis for [Sunday is available here.](SundayAnalysis.md)

# Conclusions

Based on the analysis, the bagged tree model performed better on the weekend data because the misclassification rate was about 30% and 40%. This means that 60-70% of the time the model selected the correct `shares` group. The bagged tree model did not perform as well as on the Monday-Friday data because the misclassification rate was around 50%. Based on the correlation plot used to select the variables for the collection of linear regression models, the subset of variables do not have much impact on the `shares` variable because the correlations were low. Another conclusion based on the analysis is there are more shares on the weekends because there was signifcantly more counts in the "above 1400" group on Saturday and Sunday compared to the week days. Lastly, it can be concluded from the linear regression selection that both fit stats and prediction error are needed to select a model because sometimes the adjusted R square and AIC/BIC values do not agree.