---
title: "ST558 Project 2"
author: "Noel Hilliard"
date: "July 3, 2020"
output:
  rmarkdown::github_document:
    toc: true
params:
  weekday: "monday"
---

Project Objective: The goal is to create models for predicting the `shares` variable from the dataset. You will create two models: a linear regression model and a non-linear model (each of your choice). You will use the parameter functionality of markdown to automatically generate an analysis report for each `weekday_is_*` variable (so you’ll end up with seven total outputted documents).

# Introduction

# Online News Popularity Data

```{r read in data, echo = TRUE, message = FALSE}

library(tidyverse)

newsData <-read_csv("S:/ST558/Homeworks/Project 2/ST558 Project 2/OnlineNewsPopularity.csv")

#paste column name for day of week
day <- paste0('weekday_is_', params$weekday)

#filter for day of week
weekdayData <- newsData %>% filter(!!as.name(day) == "1")
```

```{r train/test data sets, echo = TRUE}

#Create test/train data sets from filtered data set
set.seed(1)
train <- sample(1:nrow(weekdayData), size = nrow(weekdayData)*0.7)
test <- dplyr::setdiff(1:nrow(weekdayData), train)
weekdayDataTrain <- weekdayData[train, ]
weekdayDataTest <- weekdayData[test, ]
```

# Summary Statistics

# Modeling

## Ensemble Model: Bagged Tree

```{r repeated cv bag tree, echo = TRUE, message = FALSE}
library(caret)

#set.seed(1)
#trCtrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)

#bag_fit <- train(shares ~ , data = newsDataTrain, method = "treebag", 
#trControl = trCtrl, preProcess = c("center", "scale"))

#bag_fit
```











```{r rmarkdown automation, echo = FALSE}
#rmarkdown::render("S:/ST558/Homeworks/Project 2/ST558 Project 2/README.Rmd", 
 #                 output_file = "monday.md", params = list(weekday = "monday"))
```

